---
title: "Panel regression Hospitalizations"
author: "JL-Herrera"
date: "2024-10-28"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(plm)
```

Here I describe the procedure I am using to perform panel regression to my data. My objective is to see if time (week), SVI, mobility ratio, are good predictors for hospitalizations in each zip code.

### Data
This is the data I have
```{r cars}
allCounties <-read_csv("~/Documents/GitHub/MobilityAndSVIModels/dataModelSVI_MR_Hops.csv")

# Convert data to a panel data frame
data_panel <- pdata.frame(allCounties, index = c("Zip", "week"))
#Since we know that SVI impacts the effect of mobility on hospitalizations and potentially
#hospitalizations, we include an interaction term between SVI and mobility ratio
# Create interaction term in the dataset
data_panel <- data_panel %>% mutate(Interaction_MobOutSVI=SVI_State * mobRatioOut,
                                    Interaction_MobIntraSVI=SVI_State * mobRatioIntra)
#Additionally, we will initially consider the variable "week" as numeric
data_panel <- data_panel %>% mutate(week=as.numeric(week))
data_panel
```

### Models

I explored basic models, but the models I am interested in is:

SVI_State: SVI of each zip code, calculated comparing with zip codes from all TX.
mobRatioOut_mean: mobility ratio - out, this is the change of out-visits from the home zip code
SVI_mobInteraction: interaction term (SVI*Mobility), since we know, from our previous results in the paper, that SVI affects mobility
week: continuous variable to capture trends over weeks
we use the population of each zip code as offset, to get hospitalizations per unit population

#### Fixed effects model
```{r pressure, echo=FALSE}
#Fixed Effects Model : This model will include a numeric time variable to capture a 
#trend over weeks and control for unobserved heterogeneity across zip codes.
# Fixed effects model with interaction term and population as offset
fixed_model_interaction_off <- plm(Hosp_least ~ SVI_State + mobRatioOut + Interaction_MobOutSVI +
                                     mobRatioIntra + Interaction_MobIntraSVI + 
                                     week + offset(log(pop2019)), 
                               data = data_panel, 
                               model = "within")
summary(fixed_model_interaction_off)
```

#### Random effects model

```{r}
#Random Effects Model (for Comparison) : To see if a random effects model might be more 
#appropriate, you can try a random effects regression. You can later compare this to the 
#fixed effects model using a Hausman test.
# Random effects model
random_model <- plm(Hosp_least ~ SVI_State + mobRatioOut + Interaction_MobOutSVI +
                      mobRatioIntra + Interaction_MobIntraSVI +
                      week + offset(log(pop2019)), 
                    data = data_panel, 
                    model = "random")
summary(random_model)
```

#### Comparing models

To compare fixed effects and random effects models, we use the Hausman test, which is specifically designed to determine whether the unique errors (the time-invariant characteristics) are correlated with the regressors. The Hausman test compares the fixed effects and random effects models. If the test is significant, it suggests that the fixed effects model is more appropriate (since it assumes correlation between the individual effects and the regressors), while a non-significant result suggests that the random effects model is suitable.

```{r}
hausman_test <- phtest(fixed_model_interaction_off, random_model)
print(hausman_test)
```

Up to now, it looks like the random effects model is a better model. Now, I will see some diagnostic plots

#### Diagnostics

Fixed effects

```{r}
# Fitted values and residuals for random effects model
fitted_values <- fitted(fixed_model_interaction_off)
residuals <- resid(fixed_model_interaction_off)

#Residuals vs. Fitted Plot
fig1<-ggplot(data = data.frame(fitted = fitted_values, residuals = residuals), aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
#Q-Q Plot
# Create a data frame for the plot
fig2<-ggplot(data = data.frame(sample = residuals), aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "red", linetype = "dashed") +
  labs(title = "Q-Q Plot of Residuals", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()
#Residual Density Plot
fig3<-ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black") +
  geom_density(color = "red", size = 1) +
  labs(title = "Residual Density Plot", x = "Residuals", y = "Density") +
  theme_minimal()
#Scale-Location Plot
std_residuals <- residuals / sd(residuals)  # Standardized residuals
fig4<-ggplot(data = data.frame(fitted = fitted_values, std_residuals = abs(std_residuals)), aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(title = "Scale-Location Plot", x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal()

plot_grid(fig1,fig2,fig3,fig4,ncol=2)
```

Random effects

```{r}
# Fitted values and residuals for random effects model
fitted_values <- fitted(random_model)
residuals <- resid(random_model)

#Residuals vs. Fitted Plot
fig1<-ggplot(data = data.frame(fitted = fitted_values, residuals = residuals), aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
#Q-Q Plot
# Create a data frame for the plot
fig2<-ggplot(data = data.frame(sample = residuals), aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "red", linetype = "dashed") +
  labs(title = "Q-Q Plot of Residuals", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()
#Residual Density Plot
fig3<-ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black") +
  geom_density(color = "red", size = 1) +
  labs(title = "Residual Density Plot", x = "Residuals", y = "Density") +
  theme_minimal()
#Scale-Location Plot
std_residuals <- residuals / sd(residuals)  # Standardized residuals
fig4<-ggplot(data = data.frame(fitted = fitted_values, std_residuals = abs(std_residuals)), aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(title = "Scale-Location Plot", x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal()

plot_grid(fig1,fig2,fig3,fig4,ncol=2)
```

#### Fixing overdispersion

Since I see that my models has overdispersion, I try to address this by using a negative binomial model.

```{r}
library(glmmTMB)
nb_model <- glmmTMB(Hosp_least ~ SVI_State + mobRatioOut + Interaction_MobOutSVI + 
                      mobRatioIntra + Interaction_MobIntraSVI +
                      week + offset(log(pop2019)) + (1 | Zip), 
                    data = data_panel, 
                    family = nbinom2)
summary(nb_model)
```

```{r}
# Fitted values and residuals for random effects model
fitted_values <- fitted(nb_model)
residuals <- resid(nb_model)

#Residuals vs. Fitted Plot
fig1<-ggplot(data = data.frame(fitted = fitted_values, residuals = residuals), aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
#Q-Q Plot
# Create a data frame for the plot
fig2<-ggplot(data = data.frame(sample = residuals), aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "red", linetype = "dashed") +
  labs(title = "Q-Q Plot of Residuals", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()
#Residual Density Plot
fig3<-ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black") +
  geom_density(color = "red", size = 1) +
  labs(title = "Residual Density Plot", x = "Residuals", y = "Density") +
  theme_minimal()
#Scale-Location Plot
std_residuals <- residuals / sd(residuals)  # Standardized residuals
fig4<-ggplot(data = data.frame(fitted = fitted_values, std_residuals = abs(std_residuals)), aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(title = "Scale-Location Plot", x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal()

plot_grid(fig1,fig2,fig3,fig4,ncol=2)

```

#### Other idea
Now, my data has lots of zero values
```{r}
zero_count <- sum(data_panel$Hosp_least == 0)
total_count <- nrow(data_panel)
proportion_zeros <- zero_count / total_count
proportion_zeros
```

Then, one option is to try with a zero-inflated model.

```{r}
# Fit ZINB model
zinb_model <- glmmTMB(Hosp_least ~ SVI_State + mobRatioOut + Interaction_MobOutSVI +
                        mobRatioIntra + Interaction_MobIntraSVI + week + 
                        offset(log(pop2019)) + (1 | Zip),
                      data = data_panel,
                      family = nbinom2,
                      ziformula = ~ 1)  # Adjust ziformula if you want to include other predictors for inflation
summary(zinb_model)
```

```{r}
# Fitted values and residuals for random effects model
fitted_values <- fitted(zinb_model)
residuals <- resid(zinb_model)

#Residuals vs. Fitted Plot
fig1<-ggplot(data = data.frame(fitted = fitted_values, residuals = residuals), aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
#Q-Q Plot
# Create a data frame for the plot
fig2<-ggplot(data = data.frame(sample = residuals), aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "red", linetype = "dashed") +
  labs(title = "Q-Q Plot of Residuals", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()
#Residual Density Plot
fig3<-ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black") +
  geom_density(color = "red", size = 1) +
  labs(title = "Residual Density Plot", x = "Residuals", y = "Density") +
  theme_minimal()
#Scale-Location Plot
std_residuals <- residuals / sd(residuals)  # Standardized residuals
fig4<-ggplot(data = data.frame(fitted = fitted_values, std_residuals = abs(std_residuals)), aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(title = "Scale-Location Plot", x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal()

plot_grid(fig1,fig2,fig3,fig4,ncol=2)

```

But, my model shows a warning that might ocurr due to multicolinearity among my variables. I check this
```{r}
library(car)
vif_values <- vif(lm(Hosp_least ~ SVI_State + mobRatioOut + mobRatioIntra + Interaction_MobOutSVI + Interaction_MobIntraSVI + week, data = data_panel))
print(vif_values)
```

I will test by removing the interaction term

```{r}
zinb_model_WO_inter <- glmmTMB(Hosp_least ~ SVI_State + mobRatioOut + mobRatioIntra +
                                 #Interaction_MobOutSVI +Interaction_MobIntraSVI + 
                                 week + 
                        offset(log(pop2019)) + (1 | Zip),
                      data = data_panel,
                      family = nbinom2,
                      ziformula = ~ 1)  # Adjust ziformula if you want to include other predictors for inflation
summary(zinb_model_WO_inter)
```

```{r}
# Fitted values and residuals for random effects model
fitted_values <- fitted(zinb_model_WO_inter)
residuals <- resid(zinb_model_WO_inter)

#Residuals vs. Fitted Plot
fig1<-ggplot(data = data.frame(fitted = fitted_values, residuals = residuals), aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
#Q-Q Plot
# Create a data frame for the plot
fig2<-ggplot(data = data.frame(sample = residuals), aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "red", linetype = "dashed") +
  labs(title = "Q-Q Plot of Residuals", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()
#Residual Density Plot
fig3<-ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black") +
  geom_density(color = "red", size = 1) +
  labs(title = "Residual Density Plot", x = "Residuals", y = "Density") +
  theme_minimal()
#Scale-Location Plot
std_residuals <- residuals / sd(residuals)  # Standardized residuals
fig4<-ggplot(data = data.frame(fitted = fitted_values, std_residuals = abs(std_residuals)), aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(title = "Scale-Location Plot", x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal()

plot_grid(fig1,fig2,fig3,fig4,ncol=2)

```


Now, I want to compare these three models I created:

Calculate AIC for Fixed and Random Effects Model Manually
```{r}
# Extract residual sum of squares and number of parameters from the random model
residuals_fixed <- resid(fixed_model_interaction_off)
rss_fixed <- sum(residuals_fixed^2)
n_fixed <- length(residuals_fixed)
k_fixed <- length(coef(fixed_model_interaction_off))  # Number of parameters

# Calculate log-likelihood
logLik_fixed <- -n_fixed/2 * (log(2 * pi) + log(rss_fixed/n_fixed) + 1)

# Calculate AIC
aic_fixed <- -2 * logLik_fixed + 2 * k_fixed

# Extract residual sum of squares and number of parameters from the random model
residuals_random <- resid(random_model)
rss_random <- sum(residuals_random^2)
n_random <- length(residuals_random)
k_random <- length(coef(random_model))  # Number of parameters

# Calculate log-likelihood
logLik_random <- -n_random/2 * (log(2 * pi) + log(rss_random/n_random) + 1)

# Calculate AIC
aic_random <- -2 * logLik_random + 2 * k_random

# Print AIC
print(paste("AIC for Random Effects Model:", aic_random))
print(paste("AIC for Fixed Effects Model:", aic_fixed))
```

```{r}
# AIC for negative binomial and zero-inflated models
aic_nb <- AIC(nb_model)
aic_zinb <- AIC(zinb_model)
aic_zinb_WO <- AIC(zinb_model_WO_inter)

# Combine results into a data frame
aic_values <- data.frame(
  Model = c("Fixed Effects","Random Effects", "Negative Binomial", "Zero-Inflated NB", "Zero-Inflated NB - WO interaction"),
  AIC = c(aic_fixed,aic_random, aic_nb, aic_zinb,aic_zinb_WO)
)

# Print AIC values
aic_values %>% arrange(AIC)
```

I am going to use the Zero-Inflated NB model, even when NB has the lowest AIC. It makes more sense since my Hospitalization data has 36% of zeros. One criteria to pick the best model is to calculate $\Delta AIC=AIC_i - \min{AIC}$. In our case $\Delta AIC=-2$. Then, we can use Zero-Inflated NB model.

```{r}
# Fit ZINB model
zinb_model <- glmmTMB(Hosp_least ~ SVI_State + mobRatioOut + mobRatioIntra + 
                        #Interaction_MobOutSVI + Interaction_MobIntraSVI + 
                        week + offset(log(pop2019)) + (1 | Zip),
                      data = data_panel,
                      family = nbinom2,
                      ziformula = ~ 1)  # Adjust ziformula if you want to include other predictors for inflation
summary(zinb_model)
```

I will see the distribution of my variables

```{r}
data_panel %>% dplyr::select(SVI_State,mobRatioOut,mobRatioIntra,Interaction_MobOutSVI,Interaction_MobIntraSVI) %>%
  mutate(Interaction_MobOutSVI=as.numeric(Interaction_MobOutSVI)) %>%
  mutate(Interaction_MobIntraSVI=as.numeric(Interaction_MobIntraSVI)) %>%
  pivot_longer(cols = everything(),names_to = "Variable",values_to = "Value") %>%
  ggplot(aes(x=Value,fill=Variable)) + geom_histogram() + facet_wrap(~Variable,ncol=3,scales = "free_x")
```

```{r}
data_panel_std<-data_panel %>% group_by(Zip) %>% 
  mutate(
    mobRatioOut_std = (mobRatioOut - mean(mobRatioOut, na.rm = TRUE)) / sd(mobRatioOut, na.rm = TRUE),
    mobRatioIntra_std = (mobRatioIntra - mean(mobRatioIntra, na.rm = TRUE)) / sd(mobRatioIntra, na.rm = TRUE),
#    SVI_State_std = (SVI_State - mean(SVI_State, na.rm = TRUE)) / sd(SVI_State, na.rm = TRUE),
    Interaction_MobOutSVI_std = mobRatioOut_std * SVI_State,
    Interaction_MobIntraSVI_std = mobRatioIntra_std * SVI_State
)

data_panel_std %>%
  ungroup() %>%
  dplyr::select(SVI_State,mobRatioOut_std,mobRatioIntra_std,Interaction_MobOutSVI_std,Interaction_MobIntraSVI_std) %>%
  mutate(Interaction_MobOutSVI_std=as.numeric(Interaction_MobOutSVI_std)) %>%
  mutate(Interaction_MobIntraSVI_std=as.numeric(Interaction_MobIntraSVI_std)) %>%
  pivot_longer(cols = everything(),names_to = "Variable",values_to = "Value") %>%
  ggplot(aes(x=Value,fill=Variable)) + geom_histogram() + facet_wrap(~Variable,ncol=3,scales = "free_x")
```

Now, zero-Inflated NB model in standardized data (mobility and interaction term).

```{r}
# Fit ZINB model
zinb_model_std <- glmmTMB(Hosp_least ~ SVI_State + mobRatioOut_std + mobRatioIntra_std + 
#                            Interaction_MobOutSVI_std + Interaction_MobIntraSVI_std +
                          week + 
                        offset(log(pop2019)) + (1 | Zip),
                      data = data_panel_std,
                      family = nbinom2,
                      ziformula = ~ 1)  # Adjust ziformula if you want to include other predictors for inflation
summary(zinb_model_std)
```

```{r}
aic_zinb <- AIC(zinb_model_std)
aic_zinb
```

```{r}
nb_model_std <- glmmTMB(Hosp_least ~ SVI_State + mobRatioOut_std + mobRatioIntra_std + 
                          #  Interaction_MobOutSVI_std + Interaction_MobIntraSVI_std +
                      week + offset(log(pop2019)) + (1 | Zip), 
                    data = data_panel_std, 
                    family = nbinom2)
summary(nb_model_std)
```
```{r}
# Fitted values and residuals for random effects model
fitted_values <- fitted(nb_model_std)
residuals <- resid(nb_model_std)

#Residuals vs. Fitted Plot
fig1<-ggplot(data = data.frame(fitted = fitted_values, residuals = residuals), aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
#Q-Q Plot
# Create a data frame for the plot
fig2<-ggplot(data = data.frame(sample = residuals), aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "red", linetype = "dashed") +
  labs(title = "Q-Q Plot of Residuals", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()
#Residual Density Plot
fig3<-ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black") +
  geom_density(color = "red", size = 1) +
  labs(title = "Residual Density Plot", x = "Residuals", y = "Density") +
  theme_minimal()
#Scale-Location Plot
std_residuals <- residuals / sd(residuals)  # Standardized residuals
fig4<-ggplot(data = data.frame(fitted = fitted_values, std_residuals = abs(std_residuals)), aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(title = "Scale-Location Plot", x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal()

plot_grid(fig1,fig2,fig3,fig4,ncol=2)

```

Interactions between SVI and mobility variables are not significantly correlated with hospitalizations. Now, the correlations between mobility variables are negative, and I was expecting them to be positive: more mobility, more hospitalizations. 

Additionally, I am going to take into consideration that mobility and hospitalizations are lagged: movement today might be reflected in hospitalizations in later times. Then, I calculate lags of mobility to update my model

```{r}
data_panel_with_std_lags <- data_panel_std %>%
  arrange(Zip, week) %>%
  group_by(Zip) %>%
  mutate(mobRatioOut_std_lag1 = dplyr::lag(mobRatioOut_std, 1),
         mobRatioOut_std_lag2 = dplyr::lag(mobRatioOut_std, 2),
         mobRatioOut_std_lag3 = dplyr::lag(mobRatioOut_std, 3),
         mobRatioOut_std_lag4 = dplyr::lag(mobRatioOut_std, 4),
         mobRatioOut_std_lag5 = dplyr::lag(mobRatioOut_std, 5),
         mobRatioIntra_std_lag1 = dplyr::lag(mobRatioIntra_std, 1),
         mobRatioIntra_std_lag2 = dplyr::lag(mobRatioIntra_std, 2),
         mobRatioIntra_std_lag3 = dplyr::lag(mobRatioIntra_std, 3),
         mobRatioIntra_std_lag4 = dplyr::lag(mobRatioIntra_std, 4),
         mobRatioIntra_std_lag5 = dplyr::lag(mobRatioIntra_std, 5)
  ) %>%
  ungroup()
data_panel_with_std_lags
```

And now the model with everything

```{r}
nb_model_std_and_lags <- glmmTMB(Hosp_least ~ SVI_State + mobRatioOut_std + #mobRatioIntra_std + 
                          mobRatioOut_std_lag1 + 
                            mobRatioOut_std_lag2 + mobRatioOut_std_lag3 + mobRatioOut_std_lag4 + #mobRatioOut_std_lag5 +
                            SVI_State*mobRatioOut_std_lag3+
                          #mobRatioIntra_std_lag1 + mobRatioIntra_std_lag2 + mobRatioIntra_std_lag3 + mobRatioIntra_std_lag4 + mobRatioIntra_std_lag5 +
                          #Interaction_MobOutSVI_std + Interaction_MobIntraSVI_std +
                      week + offset(log(pop2019)) + (1 | Zip), 
                    data = data_panel_with_std_lags, 
                    family = nbinom2)
summary(nb_model_std_and_lags)
```

```{r}
nb_model_std_and_lags <- glmmTMB(Hosp_least ~ SVI_State + #mobRatioOut_std + #mobRatioIntra_std + 
                          mobRatioOut_std_lag1 + 
                            #mobRatioOut_std_lag2 + 
                            mobRatioOut_std_lag3 + mobRatioOut_std_lag4 + #mobRatioOut_std_lag5 +
                            SVI_State*mobRatioOut_std_lag3+
                            (1 + mobRatioOut_std_lag3 | Zip) +
                      week + offset(log(pop2019)), 
                    data = data_panel_with_std_lags, 
                    family = nbinom2)
summary(nb_model_std_and_lags)
```

```{r}
# Fitted values and residuals for random effects model
fitted_values <- fitted(nb_model_std_and_lags)
residuals <- resid(nb_model_std_and_lags)

#Residuals vs. Fitted Plot
fig1<-ggplot(data = data.frame(fitted = fitted_values, residuals = residuals), aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
#Q-Q Plot
# Create a data frame for the plot
fig2<-ggplot(data = data.frame(sample = residuals), aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "red", linetype = "dashed") +
  labs(title = "Q-Q Plot of Residuals", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()
#Residual Density Plot
fig3<-ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black") +
  geom_density(color = "red", size = 1) +
  labs(title = "Residual Density Plot", x = "Residuals", y = "Density") +
  theme_minimal()
#Scale-Location Plot
std_residuals <- residuals / sd(residuals)  # Standardized residuals
fig4<-ggplot(data = data.frame(fitted = fitted_values, std_residuals = abs(std_residuals)), aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(title = "Scale-Location Plot", x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal()

plot_grid(fig1,fig2,fig3,fig4,ncol=2)

```



```{r}
nb_model_std_and_lags_new <- glmmTMB(Hosp_least ~ SVI_State + mobRatioOut_std_lag1 + 
                            mobRatioOut_std_lag3 + mobRatioOut_std_lag4 + I(mobRatioOut_std_lag3^2) +
                            SVI_State*mobRatioOut_std_lag3+
                            (1 + mobRatioOut_std_lag3 | Zip) +
                      week + I(week^2) + offset(log(pop2019)), 
                    data = data_panel_with_std_lags, 
                    family = nbinom2)
summary(nb_model_std_and_lags_new)
```


```{r}
# Fitted values and residuals for random effects model
fitted_values <- fitted(nb_model_std_and_lags_new)
residuals <- resid(nb_model_std_and_lags_new)

#Residuals vs. Fitted Plot
fig1<-ggplot(data = data.frame(fitted = fitted_values, residuals = residuals), aes(x = fitted, y = residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted", x = "Fitted Values", y = "Residuals") +
  theme_minimal()
#Q-Q Plot
# Create a data frame for the plot
fig2<-ggplot(data = data.frame(sample = residuals), aes(sample = sample)) +
  stat_qq() +
  stat_qq_line(color = "red", linetype = "dashed") +
  labs(title = "Q-Q Plot of Residuals", x = "Theoretical Quantiles", y = "Sample Quantiles") +
  theme_minimal()
#Residual Density Plot
fig3<-ggplot(data = data.frame(residuals = residuals), aes(x = residuals)) +
  geom_histogram(aes(y = ..density..), bins = 30, fill = "skyblue", color = "black") +
  geom_density(color = "red", size = 1) +
  labs(title = "Residual Density Plot", x = "Residuals", y = "Density") +
  theme_minimal()
#Scale-Location Plot
std_residuals <- residuals / sd(residuals)  # Standardized residuals
fig4<-ggplot(data = data.frame(fitted = fitted_values, std_residuals = abs(std_residuals)), aes(x = fitted, y = std_residuals)) +
  geom_point() +
  geom_smooth(method = "loess", color = "blue", se = FALSE) +
  labs(title = "Scale-Location Plot", x = "Fitted Values", y = "Standardized Residuals") +
  theme_minimal()

plot_grid(fig1,fig2,fig3,fig4,ncol=2)

```



```{r}

# nb_model_std_and_lags_new <- glmmTMB(Hosp_least ~ SVI_State + mobRatioOut_std_lag1 + 
#                             mobRatioOut_std_lag3 + mobRatioOut_std_lag4 + I(mobRatioOut_std_lag3^2) +
#                             SVI_State*mobRatioOut_std_lag3+
#                             (1 + mobRatioOut_std_lag3 | Zip) +
#                       week + I(week^2) + offset(log(pop2019)), 
#                     data = data_panel_with_std_lags, 
#                     family = nbinom2)

# Fit ZINB model
zinb_model_std_new <- glmmTMB(Hosp_least ~ SVI_State + mobRatioOut_std_lag1 + 
                            mobRatioOut_std_lag3 + mobRatioOut_std_lag4 + I(mobRatioOut_std_lag3^2) +
                            SVI_State*mobRatioOut_std_lag3+
                            (1 + mobRatioOut_std_lag3 | Zip) +
                      week + I(week^2) + offset(log(pop2019)), 
                    data = data_panel_with_std_lags, 
                    family = nbinom2,
                      ziformula = ~ 1)  # Adjust ziformula if you want to include other predictors for inflation
summary(zinb_model_std_new)
```










